### 2017-05-22 pkg/model4you/inst/JORS/manuscript_model4you.Rnw
first example in manuscript
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1362 255f4249-2263-4417-b1ea-8498779e27a7

@@

@@

>>>>

%% packages
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning}

@@

<<setup, echo=FALSE>>=
knitr::opts_chunk$set(cache = TRUE)
@

@@

\begin{figure}
\centering

\tikzstyle{block} = [rectangle, draw,
    text width=10em, text centered, rounded corners, minimum height=4em]
\tikzstyle{exp} = [rectangle,
    text width=10em, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, ->, line width=0.2em]

\begin{tikzpicture}[node distance = 7em, auto]
  \node [block] (clin) {Clinical trial};
  \node [block, below of=clin] (pdat) {Patient and treatment information, trial outcome};
  \node [block, below of=pdat] (frst) {Model-based forest};
  \node [block, below of=frst] (pmod) {Personalised models};
  \node [block, below of=pmod] (plot) {Dependence plot};

  \node [exp, right=1em of clin] {Real world};
  \node [exp, right=1em of pdat] {Data};
  \node [exp, right=1em of frst] {Black box model};
  \node [exp, right=1em of pmod] {Interpretable model};
  \node [exp, right=1em of plot] {Visual explanation};

  \path [line] (clin) -- (pdat) node [pos=0.5, left] {data collection};
  \path [line] (pdat) -- (frst) node [pos=0.5, left] {computation of forest};
  \path [line] (frst) -- (pmod) node [pos=0.5, left] {similarity weights};
  \path [line] (pmod) -- (plot) node [pos=0.5, left] {model coefficients};
\end{tikzpicture}

\caption{Personalised models workflow: from real world to interpretable
visualisation.}
\end{figure}

<<GBSG2_prep>>=
library("model4you")
library("survival")
library("ggplot2")
library("gridExtra")
data("GBSG2", package = "TH.data")

set.seed(2017)
@

<<GBSG2_bmod, fig.height = 5, out.width = '0.7\\textwidth', fig.cap = "Base model for GBSG2 data.">>=
bmod <- survreg(Surv(time, cens) ~ horTh, data = GBSG2, model = TRUE)
survplot(bmod)

## extract coefficients and confidence intervals
coefs <- coef(bmod)
ci <- confint(bmod)
coefs <- cbind(coefs, ci)

## prepare for plotting
digits <- 2
cf <- format(round(coefs, digits), nsmall = digits)
colnams <- colnames(cf)
colnams[1] <- "theta"
grid.table(cf, cols = colnams,
           theme = ttheme_minimal(colhead = list(fg_params = list(parse = TRUE))),
           vp = viewport(x = 0, y = 0, 
                         width = 0.7, height = 0.4,
                         just = c("left", "bottom")))
@

<<GBSG2_tree, fig.width = 9, fig.cap = "Personalised model tree for GBSG2 data.">>=
tr <- pmtree(bmod)
plot(tr, terminal_panel = node_pmterminal(tr, plotfun = survplot, 
                                          confint = TRUE))
@

<<GBSG2_pmod>>=
frst <- pmforest(object = bmod, ntree = 100)

coeffun <- function(model) {
  ## model coefficients
  coefs <- c(coef(model), scale = model$scale)

  ## difference in median survival 
  p = 0.5
  coefs["median_s0"] <- qweibull(p = p, shape = 1/coefs["scale"], 
                                 scale = exp(coefs["(Intercept)"]))
  coefs["median_s1"] <- qweibull(p = p, shape = 1/coefs["scale"], 
                                 scale = exp(coefs["(Intercept)"] + coefs[2]))
  coefs["median_sdiff"] <- coefs["median_s1"] - coefs["median_s0"]

  return(coefs)
}
coefs <- pmodel(x = frst, fun = coeffun, OOB = TRUE)

## add to data and plot
dpdat <- cbind(coefs, GBSG2)
znam <- names(GBSG2)[!(names(GBSG2) %in% c("horTh", "time", "cens"))]

dplot <- function(z) {
  p <- ggplot(dpdat, aes_string(x = z, y = "median_sdiff")) 
  if(is.factor(dpdat[, z])) {
    p + geom_boxplot()
  } else {
    p + geom_point(alpha = 0.2) + 
      geom_smooth(fill = NA, method = "loess")
  }
}
dps <- lapply(znam, dplot)
do.call(grid.arrange, dps)

do.call(grid.arrange, lapply(dps, function(p) {
  dmin <- min(dpdat$median_sdiff)
  dq <- quantile(dpdat$median_sdiff, 0.99)
  p + coord_cartesian(ylim = c(dmin, dq))
  }))
@
